---
title: "Kaggle :Students Performance in Exams"
author: "pulkit saxena"
output:
  word_document: default
  html_notebook: default
---


#Research Objective 
Study the effect of level of eductaion on students academic performance.

##Research Objectives
RO1 : Does Parents education has any effect on child eductaion performance
RO2 : Does Male students achieve high academic score than females



```{r}
#install.packages("gmodels" ,dependencies = TRUE)
library(gmodels)
library(dplyr)


#Load the data 
student_performance_data <- read.csv("StudentsPerformance.csv")

summary(student_performance_data)
```


```{r}
str(student_performance_data)
```


#Exploratory analyis
No Missing Data identified . Data consists of int and categorical variables . Distribution seem to be normally 

```{r}
hist(student_performance_data$math.score)
hist(student_performance_data$reading.score)
hist(student_performance_data$writing.score)
```

Data seems to normally distributed


#Data preparation 
Create derived variables to derive academic performance score and parents educational performance

# RO1 Does Parents education has any effect on child eductaion performance
Researcher wants to know if there is any association between parents eductaion on students academic performance 

Ho: Parents education has zero effect on child eductaion performance
Ha: Parents education has effect on child eductaion performance
```{r}

#Derive student performance based on avg score

student_performance_data$avg_score <- (student_performance_data$math.score +student_performance_data$reading.score + student_performance_data$writing.score)/3

getPerformance <- function (avg_score) {
 
  result <- 'Medium'
  
  if(avg_score >= 80) {
    # print('High')
    result <- 'High'
  } else if (avg_score > 60 && avg_score <80) {
     result <- 'Medium'
   # print('Medium')
  } else {
    result <- 'Low'
  } 

  return(result)
}


student_performance_data$ac_perf <- as.factor(mapply(getPerformance,student_performance_data$avg_score))


##RO1 check for association 
CrossTable(student_performance_data$parental.level.of.education,student_performance_data$ac_perf ,chisq=TRUE )


```

#RO1 - Result Summary
Since p-vallue is significant , researcher rejects the null hypotheis 
p =  5.028598e-06 

Hence, there is association between parents eductaion and child academic performance . 



#RO2 : Does Male students achieve high academic score than females
Researcher wants to know if male students outperform female students in academic performance or vice versa

H0: µ1 >= µ2 : Male students academic performance is greater  than or equal to female students
Ha: µ1 < µ2  : Male students academic performance is less than female employees

```{r}

#RO2 : Male students achieve high academic score - check for µ using t-test

male_dataset <- select(student_performance_data,avg_score,gender) %>% filter(gender == "male")

female_dataset <- select(student_performance_data,avg_score,gender) %>% filter(gender == "female")

t.test(male_dataset$avg_score , female_dataset$avg_score, alt = "less")

```
#RO2 - Result Summary
since p-value is signifcant  1.593e-05  researcher reject the  null hypothesis . Hence Male students academic performance is less than female employees.




