---
title: "who_suicide_statistics"
author: "pulkit saxena"
output:
  word_document: default
  html_notebook: default
---
#Dataset Overview 
WHO Suicide Statistics - Basic historical (1979-2016) data by country, year and demographic groups

#Potential  Customer
Governement,NGO and insurance companies.
Governent-NOGO : Can use this data for social welfare schemes. To understand the segment which has sucidal tendency.
Insurance Companies : The life insurance suicide clause is a provision that’s in place during the first x years of the policy. Normally, when the policyholder dies, the death benefit is paid to the beneficiaries  lump-sum amount .However, if the death is a result of suicide, the insurer can refuse to pay.
Insurance companies can use this data to define the variable x - years of polciy for suicide clause based on different segments

#Research Objectives
1. Research is limited to country US
2. Identify different segments and is there any difference between the segment regarding the suicide numbers 
   across various segments

#Exploratory data anlysis 
First we will load the data and so some exploratory analysis regarding the data , variables ,outliers etc.

```{r}
library(dplyr)
testdata <- read.csv("who_suicide_statistics.csv")

str(testdata)

```
##Data description
Country , sex, age are factor variables . For segmentation purpose demographic variables sex and age are of most importance to our analysis
Response variable which is sucicides_no  has NA

```{r}
#load the data only for US
testdata_america = testdata %>%group_by(country) %>%filter(country == "United States of America")
str(testdata_america)

summary(testdata_america)

```

#US-EDA
There is missing data for suicides_no NA's   :12 . Since there are 12 records which do not have the data .

##Missing Data Handling
Omit the records since  only 2.6% of the data is missing


```{r}
testdata_america = na.omit(testdata_america)

```

#RO1 : Does Male have higher suicdal tendency as compared to females
Researcher wants to know if there are more male suicides as compared to females

H0: µ1 <= µ2 : Male suicides is  less  than or equal to female 
Ha: µ1 > µ2  : Male suicides is greater than female 

```{r}

male_dataset <- select(testdata_america,suicides_no,sex) %>% filter(sex == "male")

female_dataset <- select(testdata_america,suicides_no,sex) %>% filter(sex == "female")

t.test(male_dataset$suicides_no , female_dataset$suicides_no, alt = "greater")

boxplot(male_dataset$suicides_no,female_dataset$suicides_no , names = c("Male","Female"),col = c("blue","pink"))

```
#RO1 : Results
The p-value of the test is 2.2e-16 which is less than the significance level alpha = 0.05. We can conclude that men’s average suicide number  is significantly different from women’s suicide . Hence men have higher suicide number than females stating that male have higher suicidal tendency . Our box plot visiualisation also interpret the same result.

#RO2 Is  the male suicide on rise  ?
```{r}
library(ggplot2)
male_dataset1 <- select(testdata_america,year,suicides_no,sex)%>% filter(sex == "male") %>% group_by(year) %>% 
  summarise(suicides_no = sum(suicides_no))

male_dataset1$suicides_no <- male_dataset1$suicides_no/min(male_dataset1$suicides_no)

ggplot(male_dataset1, aes(x=year, y=suicides_no)) +
         geom_line(color = "#00AFBB", size = 1) 
     

```
#RO2 - Results
Trend graphs shows that male suicide is rising .

#RO3 - Identification of sub-segment 
Researcher wants to know which age group is more prone to suicidal tendency among males.

```{r}
levels(testdata_america$age)


```


H0: µ1 = µ2 = µ3 = µ4 = µ5 = µ6 : There is no significant difference among age groups 
Ha: µ1 ≠ µ2 ≠ µ3 ≠ µ4 ≠ µ5 ≠ µ6  : There is signifcant difference among age groups

```{r}

#anova test 
aov_model <- aov(suicides_no ~ age, data = testdata_america)
summary(aov_model)




male_dataset2 <- select(testdata_america,year,suicides_no,sex,year,age)%>% filter(sex == "male") %>%    group_by(year,age) %>%  summarise(suicides_no = sum(suicides_no))



ggplot(male_dataset2, aes(x=year, y=suicides_no,colour  = age)) +
        geom_point(color = "#00AFBB", size = 2) + geom_line(size = 4)


```
#RO3 - Results
We reject null hypothesis , since the p-value <2e-16 is significant . Hence there is significant difference among various groups .

Trend graphs shows that highest number of suicide is among the age group of 35 -54 years. Another point to note that post 2005 , there is sharp increase in suicide among the age group of 55-74 years which is the age group of retirement or old age . So following 2 sub-segments are of importantance to researcher
35-54 years male - which is the earning class as per socio-economic data 
55-74 years male - which is the retirement age.

#Recommendation 
Government and NGO are advised to do further research and understand the root causes of suicide among these 2 segments . NGO can start suicide helpline and target this segment to calm done and show empathy so as to avoid suicidal tendency. Insurance companies can use this data to update the suicide clause accordingly while issuing the insurance policy. Important issue is the sudden increase in the suicide rate among the old age segment , which needs to be further investigated by Government agencies. 
